---
title: "hw01-doug-koerber"
author: "Doug Koerber"
date: "9/19/2017"
output: html_document
---

```{r}
knitr::knit_hooks$set(inline = as.character)
load("data/nba2017-salary-points.RData")
ls()
```

### 1) Data Preprocessing
#### Salary in millions of dollars
```{r}
salary_mil = salary / 1000000
salary_mil = round(salary_mil, digits = 2)
```
#### Replacing "R" with 0
```{r}
rookie_logic <- experience == "R"
experience_int = replace(experience, rookie_logic, 0)
experience_int = as.integer(experience_int)
rm(rookie_logic)
```
#### Position Factor
```{r}
center <- position == "C"
pf <- position == "PF"
sg <- position == "SG"
pg <- position == "PG"
sf <- position == "SF"
position_fac = replace(position, center, 'center')
position_fac = replace(position_fac, sf, 'small_fwd')
position_fac = replace(position_fac, pf, 'power_fwd')
position_fac = replace(position_fac, sg, 'shoot_guard')
position_fac = replace(position_fac, pg, 'point_guard')
position_fac = factor(position_fac)
rm(center, pf, sg, pg, sf)
```

##### Position Frequencies
```{r}
table(position_fac)
```

### 2) Scatterplot of Points and Salary
```{r}
plot(points, salary_mil, cex = 1.1, col = rgb(50, 172, 150, maxColorValue = 255), xlab = 'Points', ylab = 'Salary (in millions)', main = 'Scatterplot of Points and Salary')
```

The above plot shows a weak, but noticeable, positive relationship between points and salary.  Most players are clustered below $5,000,000 and 500 points, whereas the league's "all-stars" tend to score more points and receive higher salaries.  Unfortunately, not everyone can be an "all-star."

### 3) Correlation between Points and Salary
For this exercise, _Points_ is assigned to the variable _X_, and _Salary_ is assigned to the variable _Y_.
```{r}
n = 441
xbar = (sum(points)) / n
ybar = (sum(salary_mil)) / n
var_x = (1/(n - 1)) * (sum((points - xbar)^2))
var_y = (1/(n - 1)) * (sum((salary_mil - ybar)^2))
sd_x = sqrt(var_x)
sd_y = sqrt(var_y)
cov_xy = (1/(n - 1)) * (sum((points - xbar)*(salary_mil - ybar)))
cor_xy = cov_xy / (sd_x * sd_y)
```

### 4) Simple Linear Regression
```{r}
b1 = cor_xy * (sd_y / sd_x)
b0 = ybar - (b1 * xbar)
yhat = b0 + (b1 * points)
```
#### Summary Statistics of Y-hat
```{r}
summary(yhat)
```
#### Regression Equation: Inline Code
yhat = `r round(b0, 4)` + (`r round(b1, 4)` * points)

#### b1
b1 is the effect of points on salary in millions, or the X variable on the Y variable.  Because our salary is in millions, we can see that an increase in 1 point scored would cause a `r round(b1, 4)` millions of dollars increase in salary.

#### b0
b0 is the y-intercept of our regression line.  This means that assuming a player scores 0 points in a season, their predicted salary is `r round(b0, 4)` millions of dollars.

#### Predicted Salaries
* X = 0 points:  $`r round(b0 + (b1 * 0), 4)` Million  
* X = 100 points:  $`r round(b0 + (b1 * 100), 4)` Million  
* X = 500 points:  $`r round(b0 + (b1 * 500), 4)` Million  
* X = 1000 points:  $`r round(b0 + (b1 * 1000), 4)` Million  
* X = 2000 points:  $`r round(b0 + (b1 * 2000), 4)` Million  

### 5) Plotting the Regression Line
```{r}
plot(points, salary_mil, cex = 1.1, col = rgb(50, 172, 150, maxColorValue = 255), xlab = 'Points', ylab = 'Salary (in millions)', main = 'Scatterplot of Points and Salary', abline(b0, b1, lwd = 2, col = "Blue"))
lines(lowess(points, salary_mil), col = "Red", lwd = 2)
text(2400, 19, labels = "Regression", col = "Blue")
text(2200, 28, labels = "Lowess", col = "Red")
```

### 6) Regression Residuals and Coefficient of Determination
#### Vector of Residuals
```{r}
e_salary_mil = salary_mil - yhat
summary(e_salary_mil)
```

#### Residual Sum of Squares
```{r}
RSS = sum(e_salary_mil^2)
```
RSS = `r round(RSS, 2)`

#### Total Sum of Squares
```{r}
TSS = sum((salary_mil - ybar)^2)
```
TSS = `r round(TSS, 2)`

#### R^2
```{r}
R_sq = 1 - (RSS / TSS)
```
R^2 = `r round(R_sq, 2)`

### 7) Exploring Position and Experience
#### Scatterplot of Years of Experience and Salary
```{r}
plot(experience_int, salary_mil, cex = 1.1, col = rgb(50, 172, 150, maxColorValue = 255), xlab = 'Years of Experience', ylab = 'Salary (in millions)', main = 'Scatterplot of Years of Experience and Salary')
lines(lowess(experience_int, salary_mil), col = "Red", lwd = 2)
```
The above plot shows the estimated value of _salary_ based on _years of experience_.  The red line, a _lowess_ line, is a locally-weighted regression function, meaning that it shows estimated values for salary based on some local weighting of _years of experience_.  Values closer to the point being estimated are given more weight in a _lowess regression method_.

#### 3D Scatterplot of Points, Experience, and Salary in Millions
```{r}
library('scatterplot3d')
scatterplot3d(points, experience_int, salary_mil, color = rgb(50, 172, 150, maxColorValue = 255), xlab = 'Points', ylab = 'Years of Experience', zlab = 'Salary (in Millions)', main = 'Scatterplot of Points, Experience, and Salary', cex.symbols = 1.1)
```
The above graph is a 3-dimensional visualization of _salary_, _points_, and _years of experience_.  This graph gives us 3 variables to compare with one another, rather than the 2 in a basic scatterplot.  A player's _points_ control how far right the dot is placed, their _salary_ controls how high up the dot is placed, and their _years of experience_ controls how far back the dot is placed.

#### Boxplot of Salary conditional upon Position
```{r}
boxplot(salary_mil ~ position_fac, xlab = 'Position', ylab = 'Salary (in Millions)')
```

The above chart shows different boxplots for the salaries of each position in the NBA.  The thick black line represents the median, the lower-limit of the box represents the cutoff for the 1st quartile, the upper-limit represents the cutoff for the 3rd quartile, and the whiskers show the minimum and maximum values.  The box itself represents the interquartile range.  Note: the whiskers do not cover outliers, which instead are shown as dots above the tip of the whisker.