---
title: "Warmup 2"
author: "Doug Koerber"
output: github_document
---

```{r}
load('~/Desktop/stat133/stat133-hws-fall17/warmup02/data/nba2017-salary-points.RData')
ls()
```
### Quantitative Variable: Salary
```{r}
summary(salary)

mean(salary)

sd(salary)

min(salary)

max(salary)

median(salary)

quantile(salary, probs = c(0.25, 0.5, 0.75, 1))

hist(salary)

boxplot(salary)

plot(density(salary), main = "Density of NBA Salary")
```

* What are the typical values in each vector?
    * On average, most players make about 546 points in a season, consisting of approximately 92 free throws, 152 two-pointers, and 49 three-pointers. The mean salary is $6,187,014, and the median salary is $3,500,000.
* What's the spread in each vector?
    * Most vectors have a significant positive skew. Most players in the NBA aren't significant outliers, whereas players like LeBron James and Russell Westbrook have abnormally high values for salary and points scored, respectively.  These outliers are what cause the positive skew of the histograms for our vectors.

### Qualitative Variable: Team
```{r}
table(team)
table(team)/length(team)
barplot(table(team))
```

* What things were hard, even though you saw them in class?
    * I'm still getting used to returning, for example, the name of the player with the highest salary.  The syntax of typing out brackets, parenthases, and curly brackets is still somewhat confusing.
* What was easy(-ish) even though we haven't done it in class?
    * Creating tables and then altering the table to show relative frequencies.
* What type of "errors" you struggled with (if any)?
    * I struggled a little bit with having my data loaded while knitting the file, as well ask keeping track of where things are saved/what my current working directory is.
* What are the parts you are not fully understanding?
    * I don't fully understand the flow of assignments. I know we download the data, but do we need to download the RMD file as well? How does working with an RMD affect other parts of RStudio? What if I have multiple data sets loaded at once? Also, once we are finished, submitting to bCourses and then GitHub will take some getting used to.
* What was the most time consuming part?
    * Finding summary statistics was easy, but formulating sentences to describe concisely what those statistics mean was the most time-consuming part of this assignment.
* Did you collaborate with other students? If so, with who? In what manner?
    * I didn't collaborate with other students.
* Was there any frustrating issue? (e.g. RStudio cryptic error, one or more package not playing nice)
    * No cryptic errors or frustrating issues, thankfully.


